---
title: "Trachoma serology threshold analyses"
subtitle: "Comparison of seroprevalence and SCR estimates from Bayesian and Frequentist estimators"
author: "Ben Arnold ben.arnold@ucsf.edu"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Summary

In this analysis, we compare evaluation unit (EU) level estimates of seroprevalence and the seroconversion rate (SCR) estimated with two different approaches. A Bayesian estimator using an MCMC approach was used in the primary development of the serology thresholds because developing posterior probabilities of class membership (i.e., post elimination) was a natural extension of the estimator. However, the Bayesian approach requires somewhat sophisticated programming and computing. The Frequentist GLM estimators can be implemented very easily.  

If future studies were to use thresholds developed using the Bayesian approach yet estimate seroprevalence or the SCR using a Frequentist approach, we need to provide assurance that they are consistent!


# Preamble

```{r preamble, message = FALSE}
#-----------------------------------------
# Source project config file
# and project functions
#-----------------------------------------
library(here)
source(here("R/0-config.R"))
source(here("R/0-functions.R"))
```

# Load estimates

```{r load previous estimates}
#-------------------------------
# load the frequentist eroprevalence and SCR estimates
# there is a single estimate for each evaluation unit (EU)
#-------------------------------
d_freq_sp  <- read_rds(file=here("output","trachoma-sero-thresholds-seroprev-estimates-1to5.rds"))
d_freq_scr <- read_rds(file=here("output","trachoma-sero-thresholds-scr-estimates-glm-1to5.rds"))

#-------------------------------
# load the Bayesian seroprevalence and SCR estimates
# there is a single estimate for each evaluation unit (EU)
#-------------------------------
d_bayes_sp <- read_csv(file=here("output/output_ages_1to5/bayesian_seroprev_estimates/","bayesian_seroprev_posterior_summary-1to5yo.csv"))
d_bayes_scr <- read_csv(file=here("output/output_ages_1to5/bayesian_scr_estimates","bayes_SCR_estimates_summary_1to5yo_AllEUs.csv"))

#-------------------------------
# limit the output to EUs that are included in the analysis
# add trachoma category see make_trachoma_cat() in 0-functions.R
#-------------------------------
d_bayes_sp <- d_bayes_sp %>% 
  rename(eu_name = eu) %>%
  filter(!eu_name %in% unpublished_EUs) %>%
  make_trachoma_cat(.)

d_bayes_scr <- d_bayes_scr %>% 
  rename(eu_name = eu) %>%
  filter(!eu_name %in% unpublished_EUs) %>%
  make_trachoma_cat(.)

#-------------------------------
# confirm frequentist and bayesian estimates have the same number of rows
#-------------------------------
(nrow(d_freq_sp) == nrow(d_bayes_sp))
(nrow(d_freq_scr) == nrow(d_bayes_scr))

#-------------------------------
# harmonize variable names and stack freq + bayes
#-------------------------------
# seroprevalence
d_freq_sp2 <- d_freq_sp %>%
  select(eu_name,
         trachoma_cat, 
         seroprev, seroprev_min95, seroprev_max95) %>%
  mutate(estimator = "Frequentist")

d_bayes_sp2 <- d_bayes_sp %>%
  select(eu_name,
         trachoma_cat,
         seroprev = median, 
         seroprev_min95 = lower, 
         seroprev_max95 = upper) %>%
  mutate(estimator = "Bayesian")

d_sp <- d_freq_sp2 %>% bind_rows(d_bayes_sp2)

# SCR
d_freq_scr2 <- d_freq_scr %>%
  select(eu_name,
         trachoma_cat, 
         scr, scr_min95, scr_max95) %>%
  mutate(estimator = "Frequentist")

d_bayes_scr2 <- d_bayes_scr %>%
  select(eu_name,
         trachoma_cat,
         scr = `50%`,
         scr_min95 = `2.5%`,
         scr_max95 = `97.5%`) %>%
  mutate(estimator = "Bayesian")

d_scr <- d_freq_scr2 %>% bind_rows(d_bayes_scr2)

```


# Plot seroprev estimates by EU

## forest plot with 95% CIs
```{r plot serop by EU, fig.height=10}
#-------------------------------
# plot serop and 95% CI by the two estimators
#-------------------------------
# order estimates by Bayesian SCR
d_sp_plot <- d_sp %>% mutate(eu_name = factor(eu_name, levels=d_freq_sp2$eu_name[order(d_freq_sp2$seroprev,decreasing=FALSE)])) %>%
  arrange(eu_name,estimator)

# plot
pcols <- cbpal[c(6,2)]
plot_sp <- ggplot(data=d_sp_plot,aes(x=eu_name,color=estimator)) +
  facet_grid(trachoma_cat ~. , scales = "free_y") +
  geom_errorbar(aes(ymin=seroprev_min95*100,ymax=seroprev_max95*100),
                width=0.3,
                position = position_nudge(x=c(0.2,-0.2))) +
  geom_point(aes(y=seroprev*100),
             position = position_nudge(x=c(0.2,-0.2))) +
  scale_color_manual(values=pcols) +
  scale_y_continuous(breaks=seq(0,60,by=10)) +
  labs(x="",y="Seroprevalence (%)")+
  coord_flip(ylim=c(0,60)) +
  theme_minimal() +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size=12),
    axis.title.x = element_text(size=12)
  )

```


```{r scatter plot of seroprev by EU}
#-------------------------------
# plot seroprev for the two estimators
#-------------------------------
# pivot seroprev wide for plot
d_sp_plot_wide <- d_sp_plot %>%
  select(eu_name,trachoma_cat,estimator,seroprev) %>%
  pivot_wider(id_cols=c("eu_name","trachoma_cat"),names_from=estimator,values_from=seroprev)

pcols <- c("#A42820", "cadetblue","gray95") 
plot_sp_scatter <- ggplot(data=d_sp_plot_wide, aes(fill=trachoma_cat)) +
  geom_abline(intercept=0,slope=1, color="gray50",linetype = "dashed") + 
  geom_point(aes(x=`Bayesian`*100,y=`Frequentist`*100), 
             pch=21,size=2,alpha=0.8) +
  labs(x="Seroprevalence (%)\nBayesian MCMC estimator",y="GLM estimator\nSeroprevalence (%)",tag="A") +
  scale_x_continuous(breaks=seq(0,60,by=10)) +
  scale_y_continuous(breaks=seq(0,60,by=10)) +
  scale_fill_manual(values=pcols, guide_legend(title="EU category")) +
  coord_cartesian(ylim=c(0,50),xlim=c(0,50)) +
  theme_minimal() +
  theme(axis.text = element_text(size=12),
        legend.position = "left",
        panel.grid.minor = element_blank(),
        plot.tag = element_text(face="bold"))

# add bounding box for panel B
plot_sp_scatter_with_box <- plot_sp_scatter +
  geom_rect(aes(xmin=-0.2,xmax=10.2,ymin=-0.2, ymax=10.2), color="black",fill=NA,linewidth=0.2) +
  geom_segment(aes(x=12,xend=52,y=5,yend=5), arrow = arrow(length=unit(0.3,"cm")),linewidth=0.2) 


#-------------------------------
# same plot, but limit to approaching elimination and post-elimination EUs
#-------------------------------
# zoom in to the seroprev <=18 range
plot_sp_scatter_zoom <- plot_sp_scatter +
  scale_x_continuous(breaks=seq(0,10,by=2)) +
  scale_y_continuous(breaks=seq(0,10,by=2)) +
  coord_cartesian(xlim=c(0,10),ylim=c(0,10)) +
  labs(tag=" ") +
  theme(legend.position = "none")

plot_sp_scatter_comp <- plot_sp_scatter_with_box + plot_sp_scatter_zoom + plot_layout(ncol=2,nrow=1, widths = c(1.1,1))


```

# Plot SCR estimates by EU

## forest plot with 95% CIs
```{r plot SCRs by EU, fig.height=10}
#-------------------------------
# plot SCRs and 95% CI by the two estimators
#-------------------------------
# order estimates by Bayesian SCR
d_scr_plot <- d_scr %>%
  mutate(eu_name = factor(eu_name, levels=d_freq_scr2$eu_name[order(d_freq_scr2$scr,decreasing=FALSE)])) %>%
  arrange(eu_name,estimator)

# plot
pcols <- cbpal[c(6,2)]
plot_scrs <- ggplot(data=d_scr_plot,aes(x=eu_name,color=estimator)) +
  facet_grid(trachoma_cat ~. , scales = "free_y") +
  geom_errorbar(aes(ymin=scr_min95*100,ymax=scr_max95*100),
                width=0.3,
                position = position_nudge(x=c(0.2,-0.2))) +
  geom_point(aes(y=scr*100),
             position = position_nudge(x=c(0.2,-0.2))) +
  scale_color_manual(values=pcols) +
  scale_y_continuous(breaks=seq(0,24,by=2)) +
  labs(x="",y="seroconversion rate per 100 child-years")+
  coord_flip(ylim=c(0,24)) +
  theme_minimal() +
  theme(legend.position = "top",
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size=12),
        axis.title.x = element_text(size=12))

```

## Scatter plot

```{r scatter plot of SCRs by EU}
#-------------------------------
# plot SCRs, scatter
#-------------------------------
# pivot SCRs wide for plot
d_scr_plot_wide <- d_scr_plot %>%
  select(eu_name, trachoma_cat, estimator,scr) %>%
  pivot_wider(id_cols=c("eu_name","trachoma_cat"),names_from=estimator,values_from=scr)

pcols <- c("#A42820", "cadetblue","gray95") 
plot_scr_scatter <- ggplot(data=d_scr_plot_wide, aes(fill = trachoma_cat)) +
  labs(x="SCR per 100 PY\nBayesian MCMC estimator",y="GLM estimator\nSCR per 100 PY", tag="B") +
  scale_x_continuous(breaks=seq(0,24,by=4)) +
  scale_y_continuous(breaks=seq(0,24,by=4)) +
  coord_cartesian(ylim=c(0,24),xlim=c(0,24)) +
  theme_minimal() +
  theme(axis.text = element_text(size=12),
        legend.position = "none",
        panel.grid.minor = element_blank(),
        plot.tag = element_text(face="bold"))

# add bounding box for panel B
plot_scr_scatter_with_box <- plot_scr_scatter +
  geom_rect(aes(xmin=-0.2,xmax=5.2,ymin=-0.2, ymax=5.2), color="black",fill="white",linewidth=0.2) +
  geom_segment(aes(x=6,xend=25,y=2,yend=2), arrow = arrow(length=unit(0.3,"cm")),linewidth=0.2) +

  geom_abline(intercept=0,slope=1, color="gray50",linetype = "dashed") + 
  geom_point(aes(x=`Bayesian`*100,y=`Frequentist`*100), 
             pch=21,size=2.5,alpha=0.8
             ) +
  scale_fill_manual(values=pcols, guide_legend(title="EU category"))


#-------------------------------
# same plot, but limit to approaching elimination and post-elimination EUs
#-------------------------------
# zoom in to the SCR <=5 range
plot_scr_scatter_zoom <- plot_scr_scatter +
  geom_abline(intercept=0,slope=1, color="gray50",linetype = "dashed") + 
  geom_point(aes(x=`Bayesian`*100,y=`Frequentist`*100), 
             pch=21,size=2.5,alpha=0.8
             ) +
  scale_fill_manual(values=pcols, guide_legend(title="EU category")) +
  scale_x_continuous(breaks=0:5) +
  scale_y_continuous(breaks=0:5) +
  coord_cartesian(xlim=c(0,5),ylim=c(0,5)) +
  labs(tag=" ") +
  theme(legend.position = "none")

plot_scr_scatter_comp <- plot_scr_scatter_with_box + plot_scr_scatter_zoom + plot_layout(ncol=2,nrow=1, widths = c(1.1,1))
plot_scr_scatter_comp

```


# Composite figure
```{r plot composite figure}
#-------------------------------
# make a composite figure with seroprevalence on top row and scr on the bottom row
#-------------------------------
plot_comp <- plot_sp_scatter_with_box + plot_sp_scatter_zoom + 
  plot_scr_scatter_with_box + plot_scr_scatter_zoom + 
  guide_area() + plot_annotation(caption = "PY: person-years") +
  plot_layout(ncol=2, nrow=3, widths = c(1,1), heights = c(1,1,1), guides = "collect")
  
ggsave(filename=here("output","FigS8.png"),plot_comp,device = "png",height=220,width=180, units="mm")

```


# Session Info
```{r session info}
sessionInfo()
```

